Weak interspecific interactions in a sagebrush steppe: evidence from observations, models, and experiments

### 2017-01-30 removals/manuscript/cover.tex
Final edits to ms and cover letter

Interspecific competition is arguably the primary obsession of plant community ecologists, but in a series of paper my collaborators and I have published based on long-term demographic data and multispecies population models for semiarid plant communities (Adler et al. 2010, Adler et al. 2012, Chu and Adler 2015), we found surprisingly weak interspecific interactions. These models predict little competitive release: removal of a dominant species would have little impact on the abundances of remaining species.

We conducted a five year removal experiment in a sagebrush steppe plant community to test this prediction. We found that models based on long-term observational data adequately 
predicted vital rate and population-level responses to removals for three of our four study species. For the fourth species, the model based on observational data underestimated population growth in the removal treatment, but results of additional analyses question whether this mismatch actually reflects competitive release.

Overall, the results of our experiment increase our confidence that the structure and dynamics of this plant community are determined more by variation in intraspecific interactions than by strong interspecific competition for resources. We expect this conclusion will provide food for thought for a broad audience of plant ecologists raised on a steady diet of studies emphasizing the importance of interspecific competition.

>>>>

Interspecific competition is arguably the primary obsession of plant community ecologists. However, in a series of papers based on long-term demographic data and multispecies population models for semiarid plant communities (Adler et al. 2010, Adler et al. 2012, Chu and Adler 2015), my collaborators and I found surprisingly weak interspecific interactions. These results imply little competitive release; removal of a dominant species should have little impact on the abundances of remaining species.

To test this prediction, we conducted a five year removal experiment in a sagebrush steppe plant community. We found that models based on long-term observational data adequately 
predicted vital rate and population-level responses to removals for three of our four study species. For the fourth species, the model based on observational data underestimated population growth in the removal treatment, but additional analyses suggest that this mismatch may not actually reflect competitive release.

Overall, the results of our experiment increase our confidence that the structure and dynamics of this plant community are determined more by variation in intraspecific interactions than by strong interspecific competition for resources. We expect this conclusion may surprise a broad audience of plant ecologists accustomed to studies emphasizing the importance of interspecific competition.
